\acknowledgments
For the past few years, I have studied a topic in computer science in more than a reasonable amount of detail.
This inevitably meant that I isolated myself somewhat from \emph{normal} people.
Luckily, a few people were willing to follow me in my mental endeavors.
Not only that, some even supplied valuable guidance, for which I am very grateful.

First and foremost, I want to thank one of the most sincere academics that I know, my supervisor, Leen Torenvliet.
I have never caught him thinking of The~University as a producer of results instead of as a place where humans cultivate and share knowledge.
Because of that mentality, I was allowed to use my curiosity as the most important guidance in the creation of this thesis.

Next, I am much obliged to Ronald de Wolf and Peter van Emde Boas.
Both went out of their way to see the production of this thesis through to its completion.
If it was not for Ronald, there would be so little redundancy in this thesis that its information content would resemble random noise.
His guidance made all the difference.
As for Peter, an earlier version of this thesis was, in many places, only \enquote{correct up to pesky little details}.
The fact that the current version is much improved in this regard is the result of his many useful and pleasantly specific comments.

Also outside the ILLC community, I have met people that were willing to help out where they could.
In Ralph Bottesch, I found someone equally quirky in his approach to parameterized complexity theory as myself.
I much appreciate our collaboration.
At this point, I will also mention the awesome teachers I had at the Piter Jelles gymnasium.
One of them, Kobus Sijbrandi, has even been involved with this thesis specifically, by being the first to attempt to translate the abstract into Frisian.

On a more personal note, a special \emph{thank you} goes to Jenny Batson.
When my thesis was getting the better of me and I started questioning my own sanity, she noticed.
For me, having a cup of tea with her in her office will mark the point from which things progressed positively again.

Speaking of offices and sanity, I hope my office mates, wherever they are, have still renounced office chairs in favor of exercise balls.
Giovanni Cin{\`a}, Malvin Gattinger, and Dieuwke Hupkes, I am very happy to have had you as office mates.
Likewise, I am happy to have had Frederik Lauridsen in an office nearby.
His PhD~program was just ahead of mine, so by watching Frederik closely, I often knew what awaited me.

Outside the academic environment, Martijn de~Niet has been a great help by reminding me that the modern world is shaped by human desire and perceived need.
Applying this political view to research may not be optimistic, but it did make it easier for me to untangle myself from my research emotionally.

Much of my life outside the academic environment revolved around kayaking and I am grateful for my wonderful friends at the Rotterdamsche Cano Club.
Particular thanks go to Jaap Kraaijenga for his remarkable ability to sense what is going on in peoples lives.

With regards to kayaking, I am deeply indebted to my K2~partners, Koen van~Ginkel, Arthur Hamel, Alexandra van den Elsen, and Erwin Verkleij.
Training and racing with them has given each of my years as a PhD~student its own character.
The most memorable races were team efforts, and would not have been possible without the amazing supporting job done by Ron Hoogwater and Hans Borsboom.
I am delighted to have them as my paranymphs for the defense of this thesis.

\begin{flushright}
\theauthor \\*
Leiden, December 2019
\end{flushright}
